Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-12-03 DryDep.tex
Suggestions by Frode - Part I.

@@

  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 to the nine land surface types used in the EMAP scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}

@@

From OpenIFS an accumulated surface PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be de-accumulated. Since this field is already integrated presumably over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD. Therefor, we will refer to the de-accumulated field as PPFD. An example PAR for January 2 2005 is shown in Supplement~S.5.1. De-Accumulation of hour 3 till 21, simply means

@@

The resulting PPFD fields are still accumulated over a time period of three hours and should be divided by $3$. A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations. To counter this, we decided to refrain from the division at this stage, but need to bare this in mind for later OpenIFS cycles.

@@

Due to the divergence between the implementation of the EMEP scheme and the previous Wesely scheme with respect to dry deposition velocities and surface types, it is not possible to fully disentangle and trace back differing results to a single change. Therefor, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme. A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models. For all model integrations, the meteorological reference year is 2005. This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years. After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}. As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively. During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions. Especially in Europe, this basically results in the suppression of canopy resistance in northern Scandinavia. The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}). Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}). Accidentally, we have used forcings for the year 2014 instead of 2005. Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}). As mentioned previously, the simulations initially ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$. A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).

@@

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
    \middlehline
    Wesely\_type      & x & \multicolumn{3}{c}{n/a}  & x & n/a   & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition. Surface ozone increases globally except for regions covered by tropical forest. Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to the divergence especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme. We cover this in more detail in the following section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with modeling results}
\label{subsec:model}
%
In our evaluation of results, we closely follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF~HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively separated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare also with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low absolute variance. Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics). Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $54-70\,\unit{^\circ N}$. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.

Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
\end{equation}
with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$. This way of retrospectively computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}. We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$). The resulting $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b). The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$. However, it is not clear whether diverging dry deposition velocities over ocean or deserts have the larger impact.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent: NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differentiated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not. In comparison, the available two month of the multi-model-mean display a different pattern. Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition separated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) is shown. We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference. As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface). The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme. In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum. The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH. The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively. This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
  \label{fig:mmm_drydep_hem}

>>>>

As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types. Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest. Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type. Since we have not performed a full unfolding on the data, the results should be treated with slight caution.

@@

  \caption{Mapping of land surface categories. Either land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 can be chosen for mapping to the nine land surface types used in the EMAP scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}

@@

From OpenIFS an accumulated surface PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be de-accumulated. Since this field is already integrated presumably over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD. Therefore, we will refer to the de-accumulated field as PPFD. An example PAR for January 2 2005 is shown in Supplement~S.5.1. De-Accumulation of hour 3 till 21, simply means

@@

The resulting PPFD fields are still accumulated over a time period of three hours and should be divided by $3$. A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations. To counter this, we decided to refrain from the division at this stage, but need to bear this in mind for later OpenIFS cycles.

@@

Due to the divergence between the implementation of the EMEP scheme and the previous Wesely scheme with respect to dry deposition velocities and surface types, it is not possible to fully disentangle and trace back differing results to a single change. Therefore, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme. A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models. For all model integrations, the meteorological reference year is 2005. This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years. After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}. As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively. During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions. Especially in Europe, this basically results in the suppression of canopy resistance in northern Scandinavia. The following sensitivity study therefore comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}). Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}). Accidentally, we have used emissions for the year 2014 instead of 2005. Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}). As mentioned previously, the simulations initially ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$. A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).

@@

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Emissions\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
    \middlehline
    Wesely\_type      & x & \multicolumn{3}{c}{n/a}  & x & n/a   & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition. Surface ozone increases globally except for regions covered by tropical forest. Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to the divergence especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme. We cover this in more detail in the following section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with modeling results}
\label{subsec:model}
%
In our evaluation of results, we closely follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF~HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively separated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare also with results from \citet{ACP:Luhar2017}.

Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
\end{equation}
with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$. This way of retrospectively computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}. We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low absolute variance. Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics). Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $54-70\,\unit{^\circ N}$. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.

The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b). The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of course independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$. However, it is not clear whether diverging dry deposition velocities over ocean or deserts have the larger impact.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent: NH and SH greening season. Spring and summer in the NH is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while it is absent in winter (SH summer). Spring and summer in the SH are marked by a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differentiated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not. In comparison, the available two month of the multi-model-mean display a different pattern. Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition separated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) is shown. We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference. As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface). The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme. In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum. The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH. The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively. This indicates that surface ozone in the tropics may be very sensitive to changes in emissions of ozone precursors and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
